A Large-Scale Study of the Use of Eval in R

### 2020-08-29 inst/paper/main.tex
stuff

\newcommand{\eval}{\c{eval}}
\newcommand{\parse}{\c{parse}}
\newcommand{\source}{\c{source}}
\newcommand{\local}{\c{local}}
\newcommand{\unlockBinding}{\c{unlockBinding}}
\newcommand{\substitute}{\c{substitute}}
\newcommand{\datatable}{\c{data.table}}
\newcommand{\mlogit}{\c{mlogit}}
\newcommand{\mboost}{\c{mboost}}
\newcommand{\metafor}{\c{metafor}}
\newcommand{\lavaan}{\c{lavaan}}
\newcommand{\mclust}{\c{mclust}}
\newcommand{\gamlss}{\c{gamlss}}
\newcommand{\ggproto}{\c{ggproto}}
\newcommand{\ggplot}{\c{ggplot2}}
\newcommand{\base}{\c{base}}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

\subsection{Operation Mix}
In this section, we look at the operations executed by the expressions and
symbols on evaluation. R is a lazy language, so symbols can be bound to
promises that can evaluate arbitrary code. All operations in R are function
calls, even primitive operations such as addition and assignment are treated
as calls to functions implemented in the interpreter. In our corpus, we
observe 2.9G calls to R functions of which 125M happen inside
\eval. Furthermore, unlike Javascript, R code also calls into peformant
native implementations of numerical and statistical routines. In the corpus
there are 463M calls to native code of which only 31M happens inside \eval.
Qualitatively, we observe all kinds of operations being executed inside
\eval calls, from numerical computing for statistical modeling to reflection
and metaprogramming for implementing DSLs. Even testing frameworks such
\c{genthat} heavily rely upon \eval to execute test code snippets. The core
R implementation is rife with uses of \eval for mundane things like
accessing default values of formal parameters and in statistical packages
for computation.

\subsubsection{Provenance}

In this section we turn our attention to the \AllExpressionInputEvalCallPerc
expressions passed as input to \eval. There are two ways to obtain
expressions in R.
\begin{itemize}
  \item \c{substitute} function can be used to synthesize ASTs from
    expressions by replacing symbols with their bindings in the specified
    environment.
  \item \c{expression} function can be used to create a vector of expression
    objects from text.
  \item \c{parse} function and its variants \c{str2expression} and
    \c{str2lang} can be used to turn strings or text from file into R
    expressions.
\end{itemize}

In our corpus, we observe 5.2\% cases where \eval directly evaluates the
output of \c{substitute}, 0.7\% cases where output of \c{parse} and its
variants is read directly and only 704 cases for \c{expression}. Majority of
expression objects consumed by \eval are generated by their transitive uses
by other functions.

The pattern \c{eval(parse(...))} can be used for dynamic code loading. This
forms the core of \c{source} and \c{sys.source} functions in R that are
commonly used for loading code in R files in interactive settings. We
investigated the number of cases in which the output of \c{parse} and its
variants is passed to \eval, directly or transitively by tainting their
output. This corresponds to \PercentParsedCallSites of the total \eval call
sites and \PercentParsedEvals of the eval calls. We observed that very few
of the eval call (\NbParseFilesRnd in total) consume the result of calling
\c{parse} on a file. Most of the eval calls consume the result of calling
\c{parse} on a string.  We also identified one function in core R,
\c{invokeRestartInteractively} that prompts the user for input, parses it,
and passes it to \eval.

>>>>

\newcommand{\eval}{\c{eval}}
\newcommand{\evals}{{\sf eval}s\xspace}
\newcommand{\parse}{\c{parse}}
\newcommand{\source}{\c{source}}
\newcommand{\local}{\c{local}}
\newcommand{\unlockBinding}{\c{unlockBinding}}
\newcommand{\substitute}{\c{substitute}}
\newcommand{\datatable}{\c{data.table}}
\newcommand{\mlogit}{\c{mlogit}}
\newcommand{\mboost}{\c{mboost}}
\newcommand{\metafor}{\c{metafor}}
\newcommand{\lavaan}{\c{lavaan}}
\newcommand{\mclust}{\c{mclust}}
\newcommand{\gamlss}{\c{gamlss}}
\newcommand{\ggproto}{\c{ggproto}}
\newcommand{\ggplot}{\c{ggplot2}}
\newcommand{\base}{\c{base}}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

\subsection{Operation Mix}

The JavaScript paper reported on the operation mix within \evals. This means
the proportion of read, write, and call bytecodes performed by the
JavaScript interpreter. They found that some web pages were creating objects
more frequently and writing into their fields.  In R, nearly every operation
performed in the interpreter is a function call. We observe 2.9G calls and,
of those 125M happen inside \eval (there are 463M native calls, of which
only 31M are in \eval).

{\bf there is no good reason for not counting reads and writes, we can...}

%% ...The following is fluff ...

%% Qualitatively, we observe all kinds of operations being executed inside
%% \eval calls, from numerical computing for statistical modeling to reflection
%% and metaprogramming for implementing DSLs. Even testing frameworks such
%% \c{genthat} heavily rely upon \eval to execute test code snippets. The core
%% R implementation is rife with uses of \eval for mundane things like
%% accessing default values of formal parameters and in statistical packages
%% for computation.

\subsection{Provenance}

In JavaScript, the provenance of the string passed into \eval is a matter of
concern as \eval can be used to breach the security of websites. The main
danger comes from string that were obtained from external sources. If they
are allowed to flow into an \eval, then an attacker could hijack the
program.

{\bf what have they found...}

There are various ways to obtain expressions:
\begin{itemize}
  \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
  \item \c{expression}  creates a vector of expression
    objects from text.
  \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
\end{itemize}

We observe 5.2\% cases where \eval directly evaluates the output of
\c{substitute}, 0.7\% cases where output of \c{parse} is read directly and
only 704 cases for \c{expression}. Most expressions consumed by \eval are
generated by other functions.

\c{eval(parse(...))} can be used for dynamic code loading. This forms the
core of \c{source} and \c{sys.source} functions in R that are commonly used
for loading code in R files in interactive settings. We investigated the
number of cases in which the output of \c{parse} and its variants is passed
to \eval, directly or transitively by tainting their output. This
corresponds to \PercentParsedCallSites of the total \eval call sites and
\PercentParsedEvals of the eval calls. We observed that very few of the eval
call (\NbParseFilesRnd in total) consume the result of calling \c{parse} on
a file. Most of the eval calls consume the result of calling \c{parse} on a
string.  We also identified one function in core R,
\c{invokeRestartInteractively} that prompts the user for input, parses it,
and passes it to \eval.
